---
title: "Analysis of Steam Games Dataset"
author: "Vasant Kumar Mogia"
date: "`r Sys.Date()`"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

## About this Dataset

The dataset **All Steam Spiele und deren Metadaten** is a comprehensive collection of data encompassing all games available on the Steam platform, along with their corresponding metadata. It serves as a valuable resource for researchers, developers, and gaming enthusiasts interested in exploring and analyzing the vast Steam gaming ecosystem.

### Motivations for Using the Steam Games Dataset

the Steam Games Dataset provides valuable insights into gaming trends, consumer sentiment, and game performance. It helps analyze factors like game popularity, pricing, reviews, and features, enabling better predictions, game development, and targeted marketing strategies. This dataset is essential for understanding the gaming market and improving decision-making within the industry.

### Key Features

This dataset includes the following information for each game:

-   Title
-   Release date
-   Developer and publisher details
-   Original.Price & Discounted Price 
-   All.Reviews.Summary
-   Popular.Tags
-   Supported languages
-   Minimum.Requirements and more!

### Acknowledgment

Special thanks to the owner of [this GitHub repository](https://github.com/FronkonGames/Steam-Games-Scraper) for compiling and sharing the original dataset.

### Objectives

This analysis aims to:

1.  Explore and visualize trends in game genres, ratings, and more.
2.  Identify key factors that influence game market.
3.  Examine the evolution of game releases over time.

### Basic Dataset Overview

```{r echo=TRUE}

# Load the dataset
steam_data <- read.csv('archive/dataset.csv')

# Take only the first 10,000 rows
steam_data_subset <- head(steam_data, 10000)

# Check the summary of the subset
summary(steam_data_subset)

# View column names
colnames(steam_data_subset)


```

### Load packages

```{r}
# Load all necessary libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(stringr)
library(summarytools)
library(plotly)

```

### Preprocessing:
Handle missing values and incorrect data entries.

```{r echo=TRUE}
steam_data_clean <- steam_data_subset %>%
  drop_na() %>%
  distinct() %>%
  mutate(
    Release.Date = ymd(Release.Date)
  ) %>%
  mutate(
    ReleaseYear = year(Release.Date)
  ) %>%
  filter(`Discounted.Price` > 0)

summary(steam_data_clean)
colnames(steam_data_clean)


```

### Descriptive Analysis:

```{r} 

summary(steam_data_clean)
```

### Most Common Developers and Publishers

```{r}

# Top 10 developers
top_developers <- sort(table(steam_data_subset$Developer), decreasing=TRUE)[2:8]
barplot(top_developers, main="Most Common Developers", col="lightcoral", las=2, cex.names=0.5)

# Top 10 publishers
top_publishers <- sort(table(steam_data_subset$Publisher), decreasing=TRUE)[2:8]
barplot(top_publishers, main="Most Common Publishers", col="lightblue", las=2, cex.names=0.5)

```

### Popular Tags and Game Features

```{r}
# Most popular tags 
popular_tags <- unlist(strsplit(as.character(steam_data_subset$Popular.Tags), ","))

popular_tag_table <- table(popular_tags)

top_10_tags <- head(sort(popular_tag_table, decreasing = TRUE), 10)

top_10_tags

library(ggplot2)

# Prepare the top 10 most popular tags
top_10_tags_df <- data.frame(Tag = names(top_10_tags), Count = as.numeric(top_10_tags))

# Bar plot for top 10 most popular tags
ggplot(top_10_tags_df, aes(x = reorder(Tag, Count), y = Count, fill = Tag)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Top 10 Most Popular Tags", x = "Tag", y = "Count") +
  coord_flip() +  
  scale_fill_brewer(palette = "Set3") +
  theme_minimal()

# Most popular game features
game_features <- unlist(strsplit(as.character(steam_data_subset$Game.Features), ","))

game_feature_table <- table(game_features)

top_10_features <- head(sort(game_feature_table, decreasing = TRUE), 10)

top_10_features

# Prepare the top 10 most popular game features
top_10_features_df <- data.frame(Feature = names(top_10_features), Count = as.numeric(top_10_features))

# Bar plot for top 10 most popular game features
ggplot(top_10_features_df, aes(x = reorder(Feature, Count), y = Count, fill = Feature)) +
  geom_bar(stat = "identity", color = "black") +
  labs(title = "Top 10 Most Popular Game Features", x = "Game Feature", y = "Count") +
  coord_flip() +  
  scale_fill_brewer(palette = "Set3") +
  theme_minimal()

```

### Release Date Analysis

```{r} 

library(ggplot2)
library(lubridate)

steam_data_subset %>%
  mutate(
    Release.Date = mdy(Release.Date),  
    ReleaseYear = year(Release.Date)  
  ) %>%
  ggplot(aes(x = ReleaseYear)) +
  geom_bar(fill = "steelblue") +
  labs(
    title = "Number of Games Released Each Year",
    x = "Release Year",
    y = "Number of Games"
  ) +
  theme_minimal()

```

### Games with Highest Reviews

```{r}
overwhelmingly_positive_games <- subset(steam_data_subset, All.Reviews.Summary == "Overwhelmingly Positive")

library(dplyr)

selected_columns <- overwhelmingly_positive_games %>%
  select(Title, Original.Price, Discounted.Price, Release.Date, Recent.Reviews.Summary)

head(selected_columns, 10)

library(ggplot2)

```

### Games with Most Supported Languages

```{r}
languages <- unlist(strsplit(steam_data_subset$Supported.Languages, ","))

languages <- trimws(languages)

language_counts <- table(languages)

sorted_languages <- sort(language_counts, decreasing = TRUE)

head(sorted_languages, 10)

library(ggplot2)

language_df <- data.frame(Language = names(sorted_languages), Count = as.integer(sorted_languages))

ggplot(language_df[1:10, ], aes(x = reorder(Language, -Count), y = Count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  xlab("Language") +
  ylab("Count") +
  ggtitle("Top 10 Most Supported Languages in Steam Games") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Games with Most asked Minimum.Requirements 

```{r}
min_requirements <- unlist(strsplit(as.character(steam_data_subset$Minimum.Requirements), "[,\\s]+"))

min_requirements <- trimws(min_requirements)

min_requirements_table <- table(min_requirements)

top_min_requirements <- head(sort(min_requirements_table, decreasing = TRUE), 10)

top_min_requirements
```

### Review analysis 

``` {r}
unique_review_categories <- unique(steam_data_subset$All.Reviews.Summary)
print(unique_review_categories)

library(ggplot2)

steam_data_filtered <- steam_data_subset[steam_data_subset$All.Reviews.Summary != "", ]

review_counts <- as.data.frame(table(steam_data_filtered$All.Reviews.Summary))
colnames(review_counts) <- c("ReviewCategory", "Count")

# Create a pie chart using ggplot2
ggplot(review_counts, aes(x = "", y = Count, fill = ReviewCategory)) +
  geom_bar(stat = "identity", color = "black", width = 1) +  
  coord_polar(theta = "y") + 
  labs(title = "Review Summary Distribution") +
  scale_fill_brewer(palette = "Set3") +  
  theme_void() +  
  theme(legend.title = element_blank()) 


```
